---
title: "Analysis of Temporal Autocovariance Simulation Results"
output: html_document
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=12, fig.height=8, 
               echo=TRUE, warning=FALSE, message=FALSE)
```

```{r, load-libraries}
pkgload:::load_all()
library(tidyverse)
library(viridis)
#load('../simdata/expfit-flip-covs.Rdata')
load('../simdata/expfit-fluct-covs.Rdata')
opar <- par(no.readonly=TRUE)

```

## Calculate covariances

First, we average the covariances across replicates:

```{r}

covs <- expfit_fluct_res %>%
         unnest(covs) %>% 
         # see note in temp_cov() for why these times are incremented
         mutate(t0=t0+1L, t1=t1+1L) %>%  
         group_by(L, N, Va, genlen, r, t0, t1) %>% 
         summarize(cov=mean(cov))

```

Now, we look at the covariances with the initial generations of selection, 5 to
6.

```{r}

pd5_fluct <- covs %>% filter((t1 == 5 & t1 != t0) | (t0 == 5 & t1 != t0)) %>% 
          ungroup() %>% mutate(L=factor(L), gen=ifelse(t1 > 5, t1, t0)) %>% 
          filter(N==1e3)

ggplot(pd5_fluct) + geom_point(aes(gen, cov, color=L), size=0.8) + 
  facet_grid(Va ~ genlen, scales='free_y') 


```

Subset of parameters:

```{r}

rec_params <- c(0.5, 0.1, 0.01, 1.5)
va_params <- c(0.01, 0.02, 0.05)

pd5_fluctf <- pd5_fluct %>% filter(Va %in% va_params, genlen %in% rec_params)

ggplot() +  geom_point(data=pd5_fluctf, aes(gen, cov, color=as.factor(L)), 
                       alpha=0.5, size=0.8) +
  geom_hline(yintercept=0, color='red') + 
  facet_grid(Va ~ genlen, scales='free_y')  

write_tsv(pd5_fluctf, '../data/pd5f_fluct.tsv')
devtools::use_data(pd5_fluctf, overwrite=TRUE)


```

## Additive Genetic Variance Dynamics

```{r}

vark <- expfit_fluct_res %>% unnest(stats) %>% 
          filter(Va %in% va_params, genlen %in% rec_params) %>%
          group_by(genlen, N, Va, L, gen) %>% 
          summarize(zvar=mean(zvar)) 

vark %>% ggplot(aes(gen, zvar, color=as.factor(L))) + geom_point() + facet_grid(Va ~ genlen)

```


## Predictions based on a simple extension of the directional selection model

```{r}

pred_fluctf <- vark %>% filter(N==1e3) %>% ungroup() %>% 
          filter(Va %in% va_params, genlen %in% rec_params) %>%
          mutate(N=as.integer(as.character(N)), 
                 pred=pmap_dbl(list(genlen, gen-5, zvar, N), analytic_cov)) %>%
          mutate(pred=ifelse(gen >= 15, -pred, pred))


ggplot() + geom_point(data=pd5_fluctf, aes(gen, cov, color=as.factor(L)), alpha=0.3) + 
            geom_line(data=pred_fluctf, aes(gen, pred, color=as.factor(L))) + facet_grid(Va ~ genlen)


ggplot() + geom_point(data=pd5_fluctf%>% filter(genlen == 0.1, Va == 0.05), 
                      aes(gen, cov, color=as.factor(L)), alpha=0.3) + 
            geom_line(data=pred_fluctf%>% filter(genlen == 0.1, Va == 0.05), 
                      aes(gen, pred, color=as.factor(L))) + facet_grid(Va ~ genlen)


write_tsv(pred_fluctf, '../data/pred_fluctf.tsv')
devtools::use_data(pred_fluctf, overwrite=TRUE)

```



## Cumulative Variances and Covariances


```{r, cum-mat}

covs_mat_df <- expfit_fluct_res %>% 
  mutate(cdf = map(covs_mat, sum_covmat, after=5, before=25)) %>% unnest(cdf)

covs_mat_abs_df <- expfit_fluct_res %>% 
  mutate(cdf = map(covs_mat, sum_covmat, after=5, before=25, use_abs=TRUE)) %>% unnest(cdf)


Gest  <- bind_rows(
          expfit_fluct_res %>% 
           mutate(G=map_dbl(covs_mat, G, start=5, end=25, use_abs=TRUE)) %>% 
           group_by(L, N, Va, genlen) %>% summarize(G=mean(G)) %>% mutate(abs=TRUE),

         expfit_fluct_res %>% 
           mutate(G=map_dbl(covs_mat, G, start=5, end=25, use_abs=FALSE)) %>% 
           group_by(L, N, Va, genlen) %>% summarize(G=mean(G)) %>% mutate(abs=FALSE))


ggplot(Gest %>% filter(L==500, N==1e3), aes(as.factor(genlen), G, fill=abs)) + geom_bar(stat='identity') + facet_grid(~Va)

```

```{r``}

covsmat <- bind_rows(covs_mat_df %>% filter(L==500, N==1000) %>% mutate(abs=FALSE),
            covs_mat_abs_df %>% filter(L==500, N==1000) %>% mutate(abs=TRUE)) %>%
           group_by(L, N, Va, genlen, type, abs) %>% summarize(vals=mean(vals))

shape_barplot <- function(x) {
  b <- x %>% filter(abs) %>% spread(type, vals)
  c <- x %>% filter(!abs) %>% spread(type, vals)
  mat <- full_join(b, c, by=c('genlen', 'abs')) %>%
    arrange(genlen, abs) %>% select(-abs) %>% t()
  mat[is.na(mat)] <- 0
  out <- mat[-1, ]
  colnames(out) <- mat[1,]
  out
}

shape_barplot2 <- function(x) {
  mat <- x %>% spread(type, vals) %>%  t()
  out <- mat[-1, ]
  colnames(out) <- mat[1,]
  out
}

a <- covsmat %>% filter(Va %in% va_params, genlen %in% rec_params) %>% 
  ungroup() %>% select(-L, -N) %>% group_by(Va) %>% nest() %>% mutate(barplot_data=map(data, shape_barplot))

cols <- c('red', 'green', 'blue', 'purple')
par(opar)
par(mfrow=c(1, length(va_params)),
    mar=c(0, .5, 1.5, 0),
    oma=c(3, 4, 0.5, 4.5))
ymin <- -1/20; ymax <- 0.6
axs_col <- 'gray12'; cex <- 1; lwd=1
for (i in seq_along(va_params)) {
  mat <- do.call(cbind, a %>% filter(Va == va_params[i]) %>% pull(barplot_data))
  mat[1:4, ] %>% barplot(col=cols, border=0, axes=FALSE, ylim=c(ymin, ymax))
  mat[, seq(2, ncol(mat), 2)] <- 0
  mat[4, ]  %>% barplot(col=cols[3:4], add=TRUE, border=0, axes=FALSE, ylim=c(ymin, ymax))
  abline(h=0, lty='solid', col='gray12')
  if (i == 1)
    axis(2, col.axis=axs_col, cex.axis=cex*0.8, col=axs_col, 
     lwd=lwd, tck=0.018, las=1, line=1, hadj=0.6)
}

```

## Taking Absolute Value in Cumulative Covariances 

```{r}

cumdgg <- 
  expfit_fluct_res %>% 
  filter(N==1e3, L==500) %>% 
  mutate(cdf = map(covs_mat, sum_covmat, after=5, before=25, bygen=TRUE)) %>% unnest(cdf) %>%
  filter(Va %in% va_params, genlen %in% rec_params) %>%
  group_by(gen, L, N, Va, genlen, type) %>% 
  summarize(val=mean(var)) %>%
  group_by(L, N, Va, genlen, type) %>% 
  arrange(gen) %>% mutate(cumval=cumsum(val)) %>% ungroup()


cumdgg %>% filter(genlen == 0.1, Va == 0.05) %>% ggplot() + 
  geom_bar(aes(gen, cumval/gen, fill=type), alpha=0.4, position='dodge', stat='identity') + 
  #geom_point(aes(gen, val, color=type)) + 
  facet_grid(Va ~ genlen)


cumdgg_before <- 
  expfit_fluct_res %>% 
  filter(N==1e3, L==500) %>% 
  mutate(cdf = map(covs_mat, sum_covmat, after=0, before=5, bygen=TRUE)) %>% unnest(cdf) %>%
  filter(Va %in% va_params, genlen %in% rec_params) %>%
  group_by(gen, L, N, Va, genlen, type) %>% 
  summarize(val=mean(var)) %>%
  group_by(L, N, Va, genlen, type) %>% 
  arrange(gen) %>% mutate(cumval=cumsum(val)) %>% ungroup()


cumdgg_before %>% filter(genlen == 0.1, Va == 0.05) %>% ggplot() + 
  geom_bar(aes(gen, cumval, fill=type), alpha=0.4, position='dodge', stat='identity') + 
  geom_point(aes(gen, val, color=type)) + 
  facet_grid(Va ~ genlen)

# there's a subtle off by one issue here; temp_cov() uses Δp_{t-1}.
# 4th row/col is 5th timepoint. 

cumd <- 
  expfit_fluct_res %>% 
  filter(N==1e3, L==500) %>% 
  mutate(cdf = map(covs_mat, sum_covmat, after=4, before=25, bygen=TRUE)) %>% 
  unnest(cdf) %>%
  filter(Va %in% va_params, genlen %in% rec_params) %>%
  group_by(gen, L, N, Va, genlen, type) %>% 
  summarize(val=mean(var, na.rm=TRUE)) %>%
  group_by(L, N, Va, genlen, type) %>% 
  arrange(gen) %>% mutate(val=cumsum(val)) %>% ungroup()

devtools::use_data(cumd, overwrite=TRUE)

cumd_abs <- 
  expfit_fluct_res %>% 
  filter(N==1e3, L==500) %>% 
  mutate(cdf = map(covs_mat, sum_covmat, after=4, before=25, bygen=TRUE, use_abs=TRUE)) %>% 
  unnest(cdf) %>%
  filter(Va %in% va_params, genlen %in% rec_params) %>%
  group_by(gen, L, N, Va, genlen, type) %>% 
  summarize(val=mean(var, na.rm=TRUE)) %>%
  group_by(L, N, Va, genlen, type) %>% 
  arrange(gen) %>% mutate(val=cumsum(val)) %>% ungroup()

devtools::use_data(cumd_abs, overwrite=TRUE)



```

Some EDA plots (final plot is in `analyeses/figures/fluct-plots.r`).

```{r}

cumd %>% filter(genlen == 0.1, Va == 0.05)  %>%
  mutate(gen=gen+4) %>%
  ggplot() + 
  geom_bar(aes(gen, val, fill=type), alpha=0.4, position='dodge', stat='identity') + 
  #geom_point(aes(gen, val, color=type)) + 
  facet_grid(Va ~ genlen)


cumd %>% filter(genlen == 0.1, Va == 0.05)  %>%
  mutate(gen=gen+4) %>%
  ggplot() + 
  geom_bar(aes(gen, val/gen, fill=type), alpha=0.4, position='dodge', stat='identity') + 
  #geom_point(aes(gen, val, color=type)) + 
  facet_grid(Va ~ genlen)


cumd_abs %>% filter(genlen == 0.1, Va == 0.05) %>% ggplot() + 
  geom_bar(aes(gen, val/gen, fill=type), alpha=0.4, position='dodge', stat='identity') + 
  #geom_point(aes(gen, val, color=type)) + 
  facet_grid(Va ~ genlen)


```

