---
title: "Analysis of Temporal Autocovariance Simulation Results"
output: html_document
---

```{r global_options, include=FALSE}
library(knitr)
library(rmarkdown)
opts_chunk$set(fig.width=12, fig.height=8, 
               echo=TRUE, warning=FALSE, message=FALSE)
```

```{r, load-libraries}
library(devtools)
library(tidyverse)
library(viridis)
load_all()
load('../simdata/expfit-varyl-covs.Rdata')
opar <- par(no.readonly=TRUE)

```

## Dynamics of Genetic and Offspring Number Variation 

Our theoretical models for temporal autocovariance $\text{cov}(\Delta p_t,
\Delta p_s)$ under directional selection show that a key parameter is the
additive genic variation at time $s$. With our linked regions and polygenic
selection, there isn't good theory for the dynamics of additive genic
variation through time. Here, we look at the empirical dynamics of different
types of variation (additive genetic, additive genic, only LD contributions to
additive genetic variance, and offspring variance).

First, we average across simulation replicates for the different measures of
variation (and mean breeding value and fitness):

```{r, vark}

# summarize empirical variances across sims
vark <- expfit_varyl_res %>% 
          filter(Va != 0) %>% # ignore neutral sims
          unnest(stats, genic_va, allelic_cov) %>% 
          {if (all(.$gen == .$gen1)) return(.) else stop()}  %>%
          group_by(L, N, Va, alpha, genlen, r, gen) %>% 
          summarize(kvar=mean(kvar),  # variation in offspring
                    zvar=mean(zvar),  # variation in trait breeding value
                    zbar=mean(zbar),  # mean breeding value 
                    genic_va=mean(genic_va, na.rm=TRUE),  # genic variance, Va
                    ac_var=mean(ac_var),  # allelic covariance diagonal terms
                    ac_cov=mean(ac_cov)   # allelic covariance off-diagonal terms
                    ) %>% 
          ungroup()

```


### Comparing the dynamics of additive genetic variance with the decay due to drift alone

Under the infinitesimal model, the dynamics of additive genic variation, $V_a =
2 \sum_l \alpha_l^2 p_l (1-p_l)$ are driven entirely by the decay in
heterozygosity due to drift in a finite population. Here, we compare the
empirical dynamics of additive genetic variance for the trait with the
predicted decay due to drift only, e.g., 

$$
V_A(t) = V_A(0) \left(1-\frac{1}{2N}\right)^t
$$

```{r, vark-decay}
# calculate theoretic decay under finite pop size
decay_Va <- Vectorize(function(N, t, Va) Va*(1 - 1/(2*N))^t)
theory_va <- vark %>% filter(gen == 1) %>% select(L:r, zvar) %>% 
                crossing(gen=1:50) %>% group_by(L) %>% 
                mutate(theory_Va=pmap_dbl(list(N, gen, Va), decay_Va))
ggplot() + 
  geom_point(data=vark, mapping=aes(gen, zvar, color=as.factor(L)), size=0.5) + 
  geom_line(data=theory_va, mapping=aes(gen, theory_Va, color=as.factor(L))) +
  facet_grid(Va ~ genlen, scales='free_y') + 
  geom_hline(yintercept=0)

```

This model is inappropriate (unsurprisingly) for strong selection (i.e. large
$V_A$), low recombination, and smaller number of loci contributing to the trait
(small $L$).

### The dynamics of the different components of additive genetic variance

Additive genetic variance ($V_A$) is compared of two components: (1) the
additive genic variance ($V_a = 2 \sum_l \alpha_l^2 p_l(1-p_l)$) and (2) the
contribution of allelic covariance due to LD between sites contributing to the
trait ($\sum_{i \ne j} \alpha_i \alpha_j D_{ij}$). We can write the total
additive genetic variance as:

$$
V_A = 2 \sum_l \alpha_l^2 p_l(1-p_l)  + \sum_{i \ne j} \alpha_i \alpha_j D_{ij}
$$

Now, we look at the dynamics of each component, using the empirically measured
allelic covariation (effect sizes + LD components, additive genic variation,
and additive genetic variation):

```{r, vark-dyn-comp}
# compare allelic vars/covs to genic/VA 
# we use different line types let us see despite overplotting
vark %>% select(-kvar) %>%
  filter(L==500, Va > 0.01) %>%
  mutate(ac_cov=2*ac_cov, ac_var=2*ac_var, ac_tot=ac_cov + ac_var) %>%
  gather(type, var, zvar, genic_va, ac_var, ac_cov, ac_tot) %>%
  ggplot(aes(gen, var, color=type, linetype=type)) + geom_line() + facet_grid(Va ~ genlen)

```

Let's consider just $L == 500$, looking at how the allelic variances and
covariance (a proxy for genic variance and linkage disequilibria contribution
to genetic variance) changes through time:

```{r, vark-L500-ld}
vark %>% select(L:gen, ac_var, ac_cov) %>%
  filter(L == 500) %>%
  gather(type, var, ac_var:ac_cov) %>%
  ggplot(aes(gen, var, color=type)) + geom_line() + 
  facet_grid(Va ~ genlen, scales='free_y')
```

Now looking at just $V_A == 0.05$:

```{r, vark-dyn}
vark %>% select(L:gen, ac_var, ac_cov) %>%
  gather(type, var, ac_var:ac_cov) %>%
  filter(genlen > 0, Va == 0.05) %>%
  ggplot(aes(gen, var, color=type, linetype=as.factor(L))) + geom_line() + 
  facet_grid(Va ~ genlen, scales='free_y')

```

### Approximating the change in additive genetic variance with a logistic function

Can we approximate the variance through time as a logistic function?

```{r, vark-logit}

# nest the data
dvark <- vark %>% filter(Va > 0, gen >= 5) %>% 
  mutate(t = gen-5, Va0=first(zvar)) %>% group_by(N, Va, L, genlen) %>% nest()

fit_logit_va <- function(t, va, va0, weight_gamma=0.1) {
  g <- weight_gamma
  fit <- try(nls(va ~ va0 / (1 + exp(G * (t-tinfl))) + b, data.frame(t=t, va=va), 
             start=list(G=0.4, tinfl=13, b=0.01), 
             control=nls.control(maxiter=100, warnOnly=TRUE)))
  fit
}

# apply the approximation function
dvark <- dvark %>%  mutate(va0 = map_dbl(data, ~ .$zvar[.$t == 0])) %>% 
  mutate(fit = pmap(list(data, va0), ~ fit_logit_va(.x$t, .x$zvar, .y)))

dvark %>% mutate(converged=map_lgl(fit, ~ .$convInfo$isConv)) %>% 
  filter(converged) %>%
  mutate(pred=map(fit, ~ tibble(pred_va=predict(.)))) %>% unnest(data, pred) %>% 
  filter(L == 500) %>% ggplot() + geom_point(aes(t, zvar), color='blue') + 
  geom_line(aes(t, pred_va), col='red') + facet_grid(Va ~ genlen, scales='free_y')

```

Overall the logistic approximation is not bad. Note that some cells are empty,
as the `nls()` procedure failed to converge.

## Fixations

We want to ensure we don't have an unrealistic number of fixations in our
simulations. Our population size is relatively small for our main simulations
($N=1000$).

What does the proportion of fixed SNPs look like when Va > 0 and Va = 0?

```{r, fixed-snps}

# get the generation that an allele is lost or fixed
gen_fixed_or_lost <- function(x) {
  neutfreqs <- x$neut_freqs
  neutgen <- apply(neutfreqs == 0 | neutfreqs == 1, 2, function(y) min(which(y)))
  if (!is.null(x$sel_freqs)) {
    selfreqs <- x$sel_freqs
    selgen <- apply(selfreqs == 0 | selfreqs == 1, 2, function(y) min(which(y)))
    selgen[!is.finite(selgen)]  <- NA
  } else {
    selgen <- numeric()
  }
  neutgen[!is.finite(neutgen)]  <- NA
  tibble(gen=c(selgen, neutgen),
         type=c(rep('sel', length(selgen)), rep('neut', length(neutgen))))
}


# get the generation that a mutation is fixed/lost across all sims
expfit_varyl_res <- expfit_varyl_res %>% 
  mutate(genfixed=map(res, gen_fixed_or_lost)) 

expfit_varyl_res %>% unnest(genfixed) %>%
  ggplot(aes(gen, fill=type)) + geom_density(alpha=0.3) + facet_grid(Va ~ genlen)

```

There's not a huge difference in fixation times between selected and neutral
sites. Also, note that the distribution for neutral site fixation times does
not difference much between $V_A > 0$ and $V_A = 0$ simulations. The mode over
short fixation times is due to the loss of rare alleles. For example, looking
at one simulation replicate ($V_0 = 0$):

```{r, trajs}

a <- (expfit_varyl_res %>% filter(Va ==0, genlen > 4, rep %in% 1:5))

afix <- a$res[[1]]$neut_freqs
fixgen <- apply(afix == 0 | afix == 1, 2, function(x) min(which(x)))

matplot(a$res[[1]]$neut_freqs, type='l', col=c('red', 'blue')[is.finite(fixgen) +1L])

```

From this, we see low frequency alleles are getting knocked out.

## Temporal Autocovariance Models

Here, we look at temporal autocovariance, checking our theory matches our
simulations and creating plots for the paper. First, we average the covariances
across all replicates, grouping only by the simulation parameters.

```{r, calc-covs}

covs <- expfit_varyl_res %>% 
         unnest(covs) %>% 
         # see note in temp_cov() for why these times are incremented
         mutate(t0=t0+1L, t1=t1+1L) %>%  
         group_by(L, N, Va, genlen, r, t0, t1) %>% 
         summarize(cov=mean(cov))

```

### Temporal Autocovariance with generation 5 (first generation of selection)

Our temporal autocovariances are between two generations. We choose the
generation after the onset of selection, $t=5$, as a reference generation and
look how our predicted temporal autocovariance $\text{cov}(\Delta p_5, \Delta p_s)$
changes further out, where $s$ is some other generation. We first do some EDA:

```{r, calc-covs5}

pd5 <- covs %>% filter((t1 == 5 & t1 != t0) | (t0 == 5 & t1 != t0)) %>% 
          ungroup() %>% mutate(L=factor(L), gen=ifelse(t1 > 5, t1, t0))

ggplot(pd5) + geom_point(aes(gen, cov, color=L), size=0.8) + 
  facet_grid(Va ~ genlen, scales='free_y') 

```

These are the empirical covariances. Now, we superimpose our theory on these,
plugging in different genetic variances (the empirical additive genetic
variance and the empirical additive genic variance):

```{r, covs5-pred}

# for both genic and var(z)
pred_all <- vark %>% ungroup() %>% 
          mutate(N=as.integer(as.character(N)), 
                        pred_zvar=pmap_dbl(list(genlen, gen-5, zvar, N), 
                                      analytic_cov),
                        pred_genic=pmap_dbl(list(genlen, gen-5, genic_va, N), 
                                      analytic_cov))

# just zvar (for a figure)
pred <- vark %>% ungroup() %>% 
          mutate(N=as.integer(as.character(N)), 
                        pred=pmap_dbl(list(genlen, gen-5, zvar, N), 
                                      analytic_cov))

pd5r  <- pd5 %>% filter(genlen > 0) 
predr <- pred_all %>% filter(genlen<=4, genlen > 0) %>% 
  gather(type, pred, pred_zvar, pred_genic)


ggplot() +  geom_point(data=pd5r, aes(t1, cov, color=as.factor(L)), 
                       alpha=0.5, size=0.8) +
  geom_hline(yintercept=0, color='red') + 
  geom_line(data=predr, aes(gen, pred, color=as.factor(L), linetype=type)) + 
  facet_grid(Va ~ genlen, scales='free_y')  

```

Now, we see that the real covariance is driven by the additive genetic
variation, not additive genic variation. This works well for relatively loose
linkage, but not so well for tighter linked regions.

```{r}

ggplot() +  geom_point(data=pd5r %>% filter(Va == 0.05, genlen == 1.5),
                       aes(t1, cov, color=as.factor(L)), 
                       alpha=0.5, size=0.8) +
  geom_hline(yintercept=0, color='red') + 
  geom_line(data=predr %>% filter(Va == 0.05, genlen == 1.5) %>% mutate(pred = ifelse(gen < 5, NA, pred)), 
            aes(gen, pred, color=as.factor(L), linetype=type)) + 
  facet_grid(Va ~ genlen, scales='free_y')  

```

### Using a subset of the data for the publication figures

Now, we save a few of the specific cases of these data we want for the figures,
which the `figures/ms-plots.r` will use to create them.


```{r, covs5-graphics}

rec_params <- c(0.5, 0.1, 0.01, 1.5)
va_params <- c(y0.01, 0.02, 0.05)
va_params_oom <- c(0.001, 0.01, 0.1)  # for Va over orders of magnitude 

pd5f <- pd5 %>% filter(Va %in% va_params, genlen %in% rec_params)
predf <- pred_all %>% filter(Va %in% va_params, genlen %in% rec_params) %>% 
             gather(type, pred, pred_zvar, pred_genic) %>%
             mutate(L=factor(L), cov=pred) %>% filter(L == 500)


ggplot() +  geom_point(data=pd5f, aes(t1, cov, color=as.factor(L)), 
                       alpha=0.5, size=0.8) +
  geom_hline(yintercept=0, color='red') + 
  geom_line(data=predf, aes(gen, pred, color=type))  + 
  facet_grid(Va ~ genlen, scales='free_y')  + scale_y_log10() 


# tables and Rda files for plots:
write_tsv(pd5f, '../data/pd5f.tsv')
write_tsv(predf, '../data/predf.tsv')
devtools::use_data(pd5f, overwrite=TRUE)
devtools::use_data(predf, overwrite=TRUE)

```

### Comparing fit with other initial generations

Here, we check that our theory approximately holds for a temporal
autocovariance with a timepoint during selection, in this case
$\text{cov}(\Delta p_{15}, \Delta p_s)$. Note that this is a very rough
approximation, as the LD changes *considerably* during selection and our
theoretic function `analytic_cov()` assumes levels of LD determined only
mutation and drift and thus will expectedly be quite off under cases of strong
selection where LD will have changed. Consequently the fit gets worse under
stronger selections (higher levels of target $V_A$):

```{r, covs13}

refgen <- 15

pd13 <- covs %>% filter(L == 500, (t1 > t0) & t0 == refgen) %>%
          ungroup() %>% mutate(L=factor(L), gen=t1)  

pred13 <- vark %>% ungroup() %>% 
          mutate(N=as.integer(as.character(N)), 
                        pred=pmap_dbl(list(genlen, gen-refgen, zvar, N), 
                                      analytic_cov))

ggplot() +  geom_point(data=pd13 %>% filter(Va <= 0.02), aes(gen, cov, color=as.factor(L)), 
                       alpha=0.5, size=0.8) +
  geom_hline(yintercept=0, color='red') + 
  geom_line(data=pred13 %>% filter(gen >= 13, Va <= 0.02), 
            aes(gen, pred, color=as.factor(L))) + 
  facet_grid(Va ~ genlen, scales='free_y')  


pred13f <- pred13 %>% filter(gen > 13, Va < 0.02, genlen %in% rec_params, L==500)  %>% 
  mutate(Va=as.factor(Va), genlen=as.factor(genlen), L=as.factor(L)) %>%
  rename(cov = pred) # panel_plot() requires col names to be same across data/theory
pd13f <- pd13 %>% filter(Va < 0.02, genlen %in% rec_params) %>%
  mutate(Va=as.factor(Va), genlen=as.factor(genlen), L=as.factor(L))


# tables and Rda files for plots:
write_tsv(pd13f, '../data/pd13f.tsv')
write_tsv(pred13f, '../data/pred13f.tsv')
devtools::use_data(pd13f, overwrite=TRUE)
devtools::use_data(pred13f, overwrite=TRUE)

```

In the function `simpop()`, the argument `include_ld` includes the average LD
each neutral site has with all selected sites, squared. This can be used to
demonstrate that the average LD variance decreases during selection consistent
with the fit getting worse as selection continues.

### Varying N

To validate that our model works with varying population sizes (and thus
varying initial LD levels aside from the recombination level), we look at the
covariances and their analytic predictions here. These rely on different data,
as the entire varying $L$, varying $N$ dataset is rather large and was
summarized on the server.

```{r, covs-varyn}
data(pd5f_varyn)
data(predf_varyn)

ggplot() +  geom_point(data=pd5f_varyn %>% filter(L==500),
                       aes(t1, cov, color=as.factor(N)), alpha=0.5, size=0.8) +
  geom_hline(yintercept=0, color='red') + 
  geom_line(data=predf_varyn %>% filter(L==500, type=='pred_zvar', gen > 5),
            aes(gen, pred, color=as.factor(N)))  + 
  facet_grid(Va ~ genlen, scales='free_y')  

```

## Approximating the additive genetic variance dynamics with the sum of site heterozygosity

For real data, we won't have $V_A$ (it will be a parameter to be inferred). We
will, however, have the sum of site heterozygosity in the region. Here, we
look at whether SSH decay is a good proxy for $V_A$ decay through time.

```{r, ssh}

het_ssh <- 
  expfit_varyl_res %>% unnest(hets) %>% 
  mutate(neut_ssh=neut_ssh, sel_ssh=sel_ssh) %>% 
  # gather('type', 'ssh', neut_ssh, sel_ssh) %>% 
  group_by(L, N, Va, genlen, gen)  %>%
  summarize(neut_ssh=mean(neut_ssh), sel_ssh=mean(sel_ssh), 
            sel_het = mean(sel_ssh)/mean(nsel),
            neut_het = mean(neut_ssh)/mean(nneut) )

```

How different are selected and neutral site SSHs through time?

```{r, ssh-plot}

het_ssh %>% gather('type', 'het', neut_het, sel_het) %>% 
 filter(L==500) %>%
 ggplot(aes(gen, het, color=type)) + 
  geom_line() + facet_grid(Va ~ genlen)

```

The difference is minimal. How fast is the decay?

```{r, ssh-decay}

het_ssh  <- het_ssh %>% group_by(Va, genlen, N) %>% 
  mutate(neut_ssh_delta=lead(neut_het)/neut_het)

summary(het_ssh$neut_ssh_delta)

```


Next, we join the variances (and `zbar`), with the **neutral sum of site
heterozygosity** (SSH), and compare these visually. The SSH proxy variable is:

$$
V_A(s) =   \alpha_l \sum_l 2 p_l(s) (1-p_l)(s) \\
V_A(s) =   \alpha_l SSH(s) \\
V_A(s)/V_A(t) = SSH(s)/V_A(t) \\
V_A(s)= V_A(t) SSH(s)/V_A(t) \\
$$

We calculate this proxy variance for our data:

```{r, ssh-proxy}

vark_ssh <- left_join(vark, het_ssh) 

vark_ssh0 <- vark_ssh %>% filter(gen==5) %>%
               group_by(L, N, alpha, r) %>% 
               summarize(neut_ssh0=mean(neut_ssh), 
                         sel_ssh0=mean(sel_ssh), 
                         zvar0=mean(zvar))

# this uses the VA0! Not genic, but they are equivalent under exp. fitness model
vark_ssh  <- left_join(vark_ssh, vark_ssh0) %>% 
               mutate(neut_ssh_proxy=zvar0*neut_ssh/neut_ssh0,
                      sel_ssh_proxy=zvar0*sel_ssh/sel_ssh0) %>%
               ungroup()

# long version
vark_sshd <- vark_ssh %>% 
  mutate(C2=(kvar-2)/4) %>%
  gather('var_type', 'variance', 
         C2, zvar, genic_va, neut_ssh_proxy, sel_ssh_proxy) 


vark_sshd %>%
  filter(L==500) %>% ggplot() + 
  geom_line(aes(gen, variance, color=var_type, linetype=var_type)) + 
  facet_grid(alpha ~ genlen, scales='free') + simple_theme()


vark_types <- vark_sshd %>% filter(genlen %in% rec_params,
                     Va %in% va_params, L==500) %>%
       mutate(var_type=as.factor(var_type),
              variance = ifelse(gen >= 5, variance, NA))

levels(vark_types$var_type)  <- c('offspring', 'additive genic', 'neutral SSH', 
                                  'sel SSH', 'additive genetic')

write_tsv(vark_types, '../data/expfit_vark_types.tsv')
devtools::use_data(vark_types, overwrite=TRUE)

```

### How well does SSH work instead of VA?

Here, we look at the predicted temporal autocovariance from our theory.

```{r, pred-ssh}

pred_ssh <- vark_sshd %>% ungroup() %>% 
          filter(var_type %in% c('genic_va', 'neut_ssh_proxy', 'zvar')) %>%
          filter(L==500) %>% 
          mutate(cov=pmap_dbl(list(genlen, gen-5, variance, N), 
                                      analytic_cov))

pd5r  <- pd5 %>% filter(r > 0.05, r<=0.49) %>% mutate(genlen_cM=100*genlen)
predr <- pred_ssh %>% filter(r > 0.05, r<=0.49)  %>% mutate(genlen_cM=100*genlen)


ggplot() +  geom_point(data=pd5r, aes(t1, cov, color=as.factor(L)), 
                       alpha=0.5, size=0.8) +
  geom_hline(yintercept=0, color='red') + 
  geom_line(data=predr, aes(gen, cov, linetype=as.factor(var_type))) + 
  facet_grid(Va ~ genlen, scales='free_y')  + simple_theme()
```

Save all files used for predictions/data from reference generation, for
different numeric variances:

```{r}

pd5f_ssh <- pd5 %>% filter(Va %in% va_params, genlen %in% rec_params, L==500)
predf_ssh <- pred_ssh %>% 
              filter(Va %in% va_params, genlen %in% rec_params, gen>5) %>%
              mutate(var_type=factor(var_type))

levels(predf_ssh$var_type) <- c('additive genic', 'neutral SSH', 'additive genetic')

devtools::use_data(predf_ssh, overwrite=TRUE)
devtools::use_data(pd5f_ssh, overwrite=TRUE)

write_tsv(predf_ssh, '../data/predf_ssh.tsv')
write_tsv(pd5f_ssh, '../data/pd5f_ssh.tsv')


## an alternate plot with VA varying over orders of magnitude
pd5f_ssh2 <- pd5 %>% filter(Va %in% va_params_oom, genlen %in% rec_params, 
                            as.integer(as.character(L))>0)
pd5f_ssh2$L <- droplevels(pd5f_ssh2$L)
predf_ssh2 <- pred_ssh %>% 
              filter(Va %in% va_params_oom, genlen %in% rec_params, gen>5) %>%
              mutate(var_type=factor(var_type))

levels(predf_ssh2$var_type) <- c('additive genic', 'neutral SSH', 'additive genetic')

devtools::use_data(predf_ssh2, overwrite=TRUE)
devtools::use_data(pd5f_ssh2, overwrite=TRUE)

write_tsv(predf_ssh2, '../data/predf_ssh2.tsv')
write_tsv(pd5f_ssh2, '../data/pd5f_ssh2.tsv')

```

One thing we wondered is how sensitive our predictions are to the different
levels of recombination. Below, we average the numeric variances over all
recombination levels, and show the fit is very poor. 

```{r}
# averaging across recombination regimes
# summarize variances, but averaging over recombination
vark_rec <- expfit_varyl_res %>% 
              unnest(stats, genic_va, allelic_cov) %>% 
              group_by(L, N, Va, alpha, gen) %>% 
              summarize(kvar=mean(kvar), 
                        zvar=mean(zvar), 
                        zbar=mean(zbar), 
                        genic_va=mean(va),
                        ac_var=mean(ac_var),
                        ac_cov=mean(ac_cov)) %>% 
              ungroup() %>%
              # Now, add back in region genetic lengths
              crossing(genlen=unique(expfit_varyl_res$genlen))
pred_rec <- vark_rec %>% filter(genlen < 4, genlen > 0) %>%
  mutate(N=as.integer(as.character(N)), 
         pred=pmap_dbl(list(genlen, gen-5, zvar, 2*N), analytic_cov))


ggplot() + 
  geom_point(data=pd5 %>% filter(genlen >0, genlen<4), aes(t1, cov, color=as.factor(L)), 
                       alpha=0.5, size=0.8) +
  geom_hline(yintercept=0, color='red') + 
  geom_line(data=pred_rec, aes(gen, pred, color=as.factor(L))) + 
  facet_grid(Va ~ genlen, scales='free_y') 

```

## Temporal Variances though time

Now we look at the special case of $\text{cov}(\Delta p_t, \Delta p_s)$ for
$s=t$, the variance of allele frequency change. Our predictions are in blue,
the smoothed empirical values in green, and the theoretic drift ($1/2N$) value
in red:

```{r, vars}

vars <- covs %>% filter(t1 == t0) %>%
          #filter(t0 >= 5, t1>=5) %>% 
          ungroup() %>% mutate(gen=ifelse(t1 >= 5, t1, t0)-4) %>%
          left_join(vark) %>%
          filter(L==500) %>%
          mutate(pred_var= (pmap_dbl(list(genlen, 0, zvar, N), analytic_cov) + 1/2e3 )) %>%
          mutate(Ne=1/(2*cov)) 


vars_plot <- vars %>% 
  mutate(L=as.factor(L)) %>% 
  filter(genlen > 0, Va < 0.05) 

vars_plot %>%
  ggplot(aes(gen, cov)) + geom_point() + 
  geom_line(aes(t0, pred_var), color='blue') + 
  geom_hline(yintercept=1/2e3, color='red') + 
  geom_smooth(data=vars_plot %>% filter(gen>0), se=FALSE, color='green') +
  facet_grid(Va ~ genlen, scales='free_y')

```


## Cumulative Variances and Covariances

Now, we look at the cumulative variances and covariances across many
generations.

### Calculate the variances and covariances for each allele frequency change

First, we process the covariance matrices (easier to process than long tibble versions),
summarizing into var/cov components using `sum_covmat()`:

```{r, cum-mat}

covs_mat_df <- expfit_varyl_res %>% 
  mutate(cdf = map(covs_mat, sum_covmat, after=5, before=15)) %>% unnest(cdf)

```

Now, we translate these variances into $N_e$ equivalents. Note that even in
highest rec. regions, we don't see recovery to true Ne.

```{r, cum-plot}

covs_mat_df %>% filter(type=='var') %>% 
  mutate(var=vals/(16-5), Ne=1/(2*var)) %>% 
  ggplot(aes(as.factor(genlen), Ne)) + geom_boxplot() + facet_wrap(~ Va)

```

Now, we calculate the cumulative variances/covariances and create some plots
(as well as save data for the `ms-plots.R` script to generate pub figures).

```{r, pred-cum-cov}

pred_covs <- vark %>% filter(L == 500, genlen>0, genlen<4.5) %>%
                 filter(gen >= 5, gen <= 15) %>%
                 select(Va, genlen, gen, zvar, genic_va) %>%
                 group_by(Va, genlen) %>%
                 nest() %>%
                 mutate(pred_covvar=pmap(list(genlen, data), 
                      ~ pred_covvar(.y$zvar, nrow(.y), unique(.x), 1e3))) %>%
                 unnest(pred_covvar) %>% spread(type, vals) %>% 
                 rename(pred_cov=cov, pred_var=var) 


pred_covs_ssh <- vark_ssh %>% filter(L == 500, genlen>0, genlen<4.5) %>%
                 filter(gen >= 5, gen <= 15) %>%
                 select(Va, genlen, gen, neut_ssh_proxy) %>%
                 group_by(Va, genlen) %>%
                 nest() %>%
                 mutate(pred_covvar=pmap(list(genlen, data), 
                      ~ pred_covvar(.y$neut_ssh_proxy, nrow(.y), 
                                    unique(.x), 1e3))) %>%
                 unnest(pred_covvar) %>% spread(type, vals) %>% 
                 rename(pred_cov_ssh=cov, pred_var_ssh=var) 


pred_covs_all <- pred_covs %>% 
    rename(pred_cov_zvar=pred_cov, pred_var_zvar=pred_var) %>%
    left_join(pred_covs_ssh, by=c('Va', 'genlen'))


emp_covs <- 
  covs_mat_df %>% 
  filter(L==500, genlen>0, genlen < 4.5) %>% 
  mutate(vals=vals) %>% spread(type, vals) %>%
  mutate(cov=cov+var) %>%
  gather(type, vals, cov, var)  %>%
  group_by(L, N, Va, genlen, rho, type) %>% 
  summarize(mean=mean(vals, na.rm=TRUE),
            lower=quantile(vals, 0.25, na.rm=TRUE),
            upper=quantile(vals, 0.75, na.rm=TRUE)) %>%
  gather(stat, val, mean:upper) %>%
  mutate(val=val) %>% unite(col, type, stat) %>%
  spread(col, val) 

va_params2 <- c(0.001, 0.005, 0.01, 0.05, 0.08)

cumcov_plot_df <- emp_covs %>% left_join(pred_covs)  %>% 
  filter(genlen %in% rec_params, Va %in% va_params2) 
 
cumcov_plot_all_nofilt_df <- emp_covs %>% left_join(pred_covs_all)
 
cumcov_plot_all_df <- emp_covs %>% left_join(pred_covs_all)  %>% 
  filter(genlen %in% rec_params, Va %in% va_params2) 
 
cumcov_panels(cumcov_plot_df, TRUE,  ylab="var$(p_{t} - p_{0})$", 
             lx=1.8, ly=0.5, lwd=1.2, ymax=0.5)

cumcov_panels(cumcov_plot_all_df, TRUE,  ylab="var$(p_{t} - p_{0})$", 
             lx=1.8, ly=0.5, lwd=1.2, ymax=0.5, basic=FALSE)
 
# tables and Rda files for plots:
write_tsv(cumcov_plot_df, '../data/cumcov.tsv')
devtools::use_data(cumcov_plot_df, overwrite=TRUE)

write_tsv(pred_covs_all, '../data/cumcov_all.tsv')
devtools::use_data(cumcov_plot_all_df, overwrite=TRUE)

# unfiltered version
devtools::use_data(cumcov_plot_all_nofilt_df, overwrite=TRUE)

```


## Compound Parameter 

Now, we look at the how much of the parameter space of the simulations can be
understood with a single compound parameter, $V_A/R$.

```{r, compound}

pd5 %>% mutate(va_r=Va/genlen, gen=as.factor(t1)) %>% 
  filter(t0 == 5) %>% 
  ggplot(aes(va_r, cov, color=gen)) + geom_point() + 
  geom_smooth(se=FALSE) + scale_x_log10() + ylim(-0.01, 0.01)


gls <- unique(pd5$genlen)
vas <- unique(pd5$Va)
covs <- 10^{-c(-5:-1)}


vard <- crossing(Va=vas,
         genlen=gls,
         gen=c(2*(1:24))) %>% 
  filter(Va != 0, genlen != 0) %>%
  mutate(cov=pmap_dbl(list(genlen, Va, gen), 
                      ~ analytic_cov(..1, ..3, ..2, 1e3)),
         va_r=Va/genlen)

# for only cov(Δp_5, Δp_s)
pd5_vard <- pd5 %>% mutate(va_r=Va/genlen, gen=as.factor(t1)) %>% 
  filter(t0 == 5, Va > 0) %>% 
  mutate(covbin=cut_number(cov, 10)) 

## EDA
pd5_vard %>% filter(t1 < 30) %>% ggplot(aes(va_r, cov, color=gen)) + 
  geom_jitter()  + scale_x_log10() + geom_smooth() + ylim(-0.005, 0.01) 


# tables and Rda files for plots:
write_tsv(pd5_vard, '../data/pd5_vard.tsv')
devtools::use_data(pd5_vard, overwrite=TRUE)


```

## Single Replicate

Now, we look at the noise in a single replicate.

```{r, single-replicate}


est_Va_neutR2 <- function(covs, r, N, t=5, s=6) {
  emp_cov <- covs %>% filter(t0 == t, t1 == s) %>% pull(cov)
  2*emp_cov/assoc_int(s-t, r, N) 
}

get_va <- function(x, t=5) x %>% filter(gen == t) %>% pull(zvar)


expfit_varyl_res <- expfit_varyl_res %>% filter(genlen>0) %>%
      mutate(empva5_6=pmap_dbl(list(covs, genlen, N), est_Va_neutR2)) %>%
      mutate(va = map_dbl(stats, get_va))

expfit_varyl_res %>% filter(Va > 0, va < 0.2) %>%
ggplot()+ geom_point(aes(va, empva5_6)) +
  geom_abline(slope=1, intercept=0, color='cornflowerblue', size=1) +
  geom_smooth(aes(va, empva5_6), method='lm')


expfit_varyl_res %>% filter(Va > 0, va < 0.2) %>%
mutate(mse=(empva5_6 - va)^2) %>%
ggplot(aes(L, mse))+ geom_point() +
  geom_smooth(method='lm')


expfit_varyl_res %>% filter(Va > 0, va < 0.2) %>%
mutate(mse=(empva5_6 - va)^2) %>%
ggplot(aes(Va, mse))+ geom_point() +
  geom_smooth(method='lm')


expfit_varyl_res %>% filter(Va > 0) %>%
ggplot(aes(va, empva5_6, color=as.factor(genlen))) + geom_point() + 
  facet_grid(~ Va, scales='free')+ 
  geom_abline(slope=1, intercept=0, color='cornflowerblue', size=1) +
  geom_smooth(method='lm')



```

## Method of Moments

Here, we use the generalized method of moments to estimate $V_a(1)$ and $N$.
Essentially, by equating the empirical 2nd moments and cross moments for all
timepoints to their theoretic counterparts, we have an overdetermined system of
equations. We can solve this using least squares, giving us estimates of
$V_a(1)$ and $N$.

```{r, mom}

# reshape the covariance matrices for linear model 
# cached method of moments ("mom") fits
mom_fits_data <- '../data/mom_fits.rda'
if (!file.exists(mom_fits_data)) {
  mom_fits <- expfit_varyl_res %>% 
    filter(rep %in% 1:20) %>%
    mutate(mom_df = pmap(list(covs_mat, genlen, N, hets, res), 
                         ~ make_neutld_df(..1, ..2, ..3, ssh=..4, pos=..5$pos, trange=c(5, 10)))) 
    mom_fits <- mom_fits %>% mutate(lsfit = map(mom_df, fit_mom)) %>%
                mutate(params=map(lsfit, fit2params))
  devtools::use_data(mom_fits, overwrite=TRUE)
} else {
  data(mom_fits)
}

# create the dataframe for final publication figures
mom_fitsd <- mom_fits %>% unnest(params)  %>% 
  #filter(rep %in% 1:3) %>%
  mutate(emp_va=map_dbl(stats, ~ .$zvar[5])) %>%
  select(rep, L:genlen, Nest=N_est, va0_est, emp_va) %>%
  filter(genlen %in% rec_params) %>%
  mutate(genlen=as.factor(as.character(genlen)))



mom_fits %>% filter(Va != 0) %>% 
  filter(rep %in% 1:3) %>%
  mutate(emp_genic_va=map_dbl(genic_va, ~ .$genic_va[5])) %>%
  unnest(params) %>%
  ggplot(aes(emp_genic_va, va0_est, color=as.factor(genlen))) + geom_point() + 
  geom_smooth(method='lm', color='red') +
  geom_abline(yintercept=0, slope=1, color='blue') + 
  scale_x_log10() + scale_y_log10()


mom_fits %>% filter(Va != 0) %>% 
  mutate(emp_genic_va=map_dbl(genic_va, ~ .$genic_va[5])) %>%
  unnest(params) %>%
  ggplot(aes(N, N_est, color=as.factor(genlen))) + geom_boxplot() + 
  geom_smooth(method='lm', color='red') +
  geom_hline(yintercept=1e3, color='blue') + 
  scale_x_log10() + scale_y_log10()

estimate_Ne <- function(x, t0, t1) {
  # a simple Ne estimator, Ne = t/2*F
  # where F = var(p1 - p0) / ave(het1, het0)
  x <- swap_alleles(x)
  x[x == 0 | x == 1] <- NA
  het <- x * (1-x)
  F <- var(x[t1, ] - x[t0, ], na.rm=TRUE) / mean((het[t0, ] + het[t1, ]) / 2, na.rm=TRUE)
  - (t1 - t0) / (2 * log(1-F))
}


temporal_Ne_ests <- mom_fits %>% filter(Va != 0) %>%
  mutate(Ne_est = map_dbl(res, ~ estimate_Ne(.$neut_freqs, t0=5, t1=10))) %>%
  group_by(N, Va, genlen) %>% summarize(Ne_est = mean(Ne_est, na.rm=TRUE))  %>%
  mutate(genlen=as.factor(genlen))



mom_fits %>% left_join(temporal_Ne_ests) %>% 
  filter(Va !=0) %>%
  mutate(emp_genic_va=map_dbl(genic_va, ~ .$genic_va[5])) %>%
  unnest(params) %>%
  mutate(genlen_label = as.factor(genlen), Va=as.factor(Va)) %>%
  ggplot(aes(genlen_label, N_est, fill=Va)) + geom_boxplot() + 
  geom_point(aes(genlen_label, Ne_est, color=Va)) + 
  geom_hline(yintercept=1e3, color='blue') + 
  scale_y_log10() 

# combine this into the dataframe for publication figures
mom_fitsd <- mom_fitsd %>% left_join(temporal_Ne_ests) %>%
               mutate(genlen=as.factor(genlen))

# save results for graphics
devtools::use_data(mom_fitsd, overwrite=TRUE)
write_tsv(mom_fitsd, '../data/mom_fits.tsv')

```

## Fraction of Variance due to Linked Selection

Here, we look at the total fraction of selection due to linked selection (e.g.
the covariances):

```{r, frac-var-ls}

calc_total_var <- function(res, start, end) total_var(res$neut_freqs, start, end)

mom_fits_g <- expfit_varyl_res %>%
  # note that since Va = 0 cases have L == 0 by default, we 
  # need to include it here to get Va = 0 case
  filter(L == 500 | L == 0) %>%
  mutate(before_sel_var=map_dbl(res, calc_total_var, start=1, end=4)) %>%
  mutate(after_sel_var=map_dbl(res, calc_total_var, start=5, end=10)) %>%
  mutate(G_before=map_dbl(covs_mat, G, start=1, end=4)) %>%
  mutate(G_after=map_dbl(covs_mat, G, start=5, end=10)) 

devtools::use_data(mom_fits_g, overwrite=TRUE)

ggplot(mom_fits_g, aes(x=as.factor(Va), G_before, fill=as.factor(genlen))) + 
  geom_boxplot(position='dodge') 

ggplot(mom_fits_g, aes(x=as.factor(Va), G_after, fill=as.factor(genlen))) + 
  geom_boxplot(position='dodge') 

# Gprime
mom_fits_gp <- mom_fits %>% 
  # note that since Va = 0 cases have L == 0 by default, we 
  # need to include it here to get Va = 0 case
  filter(L == 500 | L == 0) %>%
  mutate(before_sel_var=map_dbl(res, calc_total_var, start=1, end=4)) %>%
  mutate(after_sel_var=map_dbl(res, calc_total_var, start=5, end=10)) %>%
  unnest(params) %>% 
  mutate(Gp_before = map2_dbl(before_sel_var, N_est, Gprime)) %>%
  mutate(Gp_after = map2_dbl(after_sel_var, N_est, Gprime))

devtools::use_data(mom_fits_gp, overwrite=TRUE)

ggplot(mom_fits_gp, aes(x=as.factor(Va), Gp_after, fill=as.factor(genlen))) + 
  geom_boxplot(position='dodge') 

ggplot(mom_fits_gp, aes(x=as.factor(Va), Gp_before, fill=as.factor(genlen))) + 
  geom_boxplot(position='dodge') 

```

## SFS under Selection

A quick look at what the SFS looks like under polygenic selection. Overall, we
don't see a large difference.

```{r}
freq_spec <- function(res, gen=50, nbins=1000) {
  effects <- res$effects
  neut_freqs <- res$neut_freqs
  num_neut_sites <- sum(is.na(effects))
  nvars <- neut_freqs[gen, ]
  freqs <- nvars[(nvars > 0) & (nvars < 1)]
  segvars_hist = cut(freqs, breaks=1/seq(0, nbins))
  tbl = table(segvars_hist)
  tibble(bin=1/seq(1, nbins), freqs=tbl)
}

sfs <- function(res, N, gen=5) {
  effects <- res$effects
  neut_freqs <- res$neut_freqs
  num_neut_sites <- sum(is.na(effects))
  nvars <- 2*N*neut_freqs[gen, ]
  segvars <- nvars[(nvars > 0) & (nvars < 2*N)]
  segvars_hist = cut(segvars, breaks=seq(0, (2*N)-1))
  tibble(freq=seq(1, 2*N -1), nsites=table(segvars_hist))
}


sfs <- function(res, N, gen=5, p0=NULL, eps=0.05) {
  effects <- res$effects
  neut_freqs <- res$neut_freqs
  num_neut_sites <- sum(is.na(effects))
  if (!is.null(p0)) {
    # use only loci within eps of p0 the first generation
    keep_loci <- abs(neut_freqs[0, ] - p0) < eps
    nvars <- 2*N*neut_freqs[gen, keep_loci]
  } else {
    nvars <- 2*N*neut_freqs[gen, ]
  }
  segvars <- nvars[(nvars > 0) & (nvars < 2*N)]
  segvars_hist = cut(segvars, breaks=seq(0, (2*N)-1))
  tibble(freq=seq(1, 2*N -1), nsites=table(segvars_hist))
}


a4 <- expfit_varyl_res %>% mutate(sfs=pmap(list(res, N), sfs, gen=4)) %>%
  unnest(sfs) %>%
  group_by(Va, genlen, L, freq) %>% summarize(nsites=mean(nsites)) %>%
  group_by(Va, genlen, L) %>% nest() %>%
     mutate(norm=map(data, function(x) tibble(freq=x$freq, nsites=x$nsites/sum(x$nsites)))) %>%
     unnest(norm)


a1 <- expfit_varyl_res %>% mutate(sfs=pmap(list(res, N), sfs, gen=1)) %>%
  unnest(sfs) %>%
  group_by(Va, genlen, L, freq) %>% summarize(nsites=mean(nsites)) %>%
  group_by(Va, genlen, L) %>% nest() %>%
     mutate(norm=map(data, function(x) tibble(freq=x$freq, nsites=x$nsites/sum(x$nsites)))) %>%
     unnest(norm)

b6 <- expfit_varyl_res %>% mutate(sfs=pmap(list(res, N), sfs, gen=6)) %>%
  unnest(sfs) %>%
  group_by(Va, genlen, L, freq) %>% summarize(nsites=mean(nsites)) %>%
  group_by(Va, genlen, L) %>% nest() %>%
     mutate(norm=map(data, function(x) tibble(freq=x$freq, nsites=x$nsites/sum(x$nsites)))) %>%
     unnest(norm)

b10 <- expfit_varyl_res %>% mutate(sfs=pmap(list(res, N), sfs, gen=10)) %>%
  unnest(sfs) %>%
  group_by(Va, genlen, L, freq) %>% summarize(nsites=mean(nsites)) %>%
  group_by(Va, genlen, L) %>% nest() %>%
     mutate(norm=map(data, function(x) tibble(freq=x$freq, nsites=x$nsites/sum(x$nsites)))) %>%
     unnest(norm)

#  Same with intermediate starting freqs
a4_int <- expfit_varyl_res %>% mutate(sfs=pmap(list(res, N), sfs, gen=4, p0=0.5)) %>%
  unnest(sfs) %>%
  group_by(Va, genlen, L, freq) %>% summarize(nsites=mean(nsites)) %>%
  group_by(Va, genlen, L) %>% nest() %>%
     mutate(norm=map(data, function(x) tibble(freq=x$freq, nsites=x$nsites/sum(x$nsites)))) %>%
     unnest(norm)


a1_int <- expfit_varyl_res %>% mutate(sfs=pmap(list(res, N), sfs, gen=1, p0=0.5)) %>%
  unnest(sfs) %>%
  group_by(Va, genlen, L, freq) %>% summarize(nsites=mean(nsites)) %>%
  group_by(Va, genlen, L) %>% nest() %>%
     mutate(norm=map(data, function(x) tibble(freq=x$freq, nsites=x$nsites/sum(x$nsites)))) %>%
     unnest(norm)

b6_int <- expfit_varyl_res %>% mutate(sfs=pmap(list(res, N), sfs, gen=6, p0=0.5)) %>%
  unnest(sfs) %>%
  group_by(Va, genlen, L, freq) %>% summarize(nsites=mean(nsites)) %>%
  group_by(Va, genlen, L) %>% nest() %>%
     mutate(norm=map(data, function(x) tibble(freq=x$freq, nsites=x$nsites/sum(x$nsites)))) %>%
     unnest(norm)

b10_int <- expfit_varyl_res %>% mutate(sfs=pmap(list(res, N), sfs, gen=10, p0=0.5)) %>%
  unnest(sfs) %>%
  group_by(Va, genlen, L, freq) %>% summarize(nsites=mean(nsites)) %>%
  group_by(Va, genlen, L) %>% nest() %>%
     mutate(norm=map(data, function(x) tibble(freq=x$freq, nsites=x$nsites/sum(x$nsites)))) %>%
     unnest(norm)



dd_int <- bind_rows(a1_int %>% mutate(type='pre-seletion-1'),
                    a4_int %>% mutate(type='pre-seletion-4'),
                    b6_int %>% mutate(type='seletion-6'),
                    b10_int %>% mutate(type='seletion-10'))



dd <- bind_rows(a1 %>% mutate(type='pre-seletion-1'),
                a4 %>% mutate(type='pre-seletion-4'),
                b6 %>% mutate(type='seletion-6'),
                b10 %>% mutate(type='seletion-10'))

dd_int %>% filter(L==50, genlen %in% c(0.01, 0.1, 0.5, 1.5), Va %in% c(0.01, 0.02, 0.05, 0.1)) %>%  
  mutate(Va=as.factor(Va)) %>%
  filter(freq > 3) %>% 
  ggplot() + 
  geom_smooth(aes(freq, nsites, color=type, group=type), 
              method='loess', se=FALSE, span=1/50) + 
  scale_x_log10() + scale_y_log10() +
  facet_grid(Va ~ genlen) +
  ylab('fraction of segregating sites')  +
  xlab('frequency')


dd %>% filter(L==50, genlen %in% c(0.01, 0.1, 0.5, 1.5), Va %in% c(0.01, 0.02, 0.05, 0.1)) %>%  
  mutate(Va=as.factor(Va)) %>%
  #filter(freq > 3) %>% 
  ggplot() + 
  geom_smooth(aes(freq, nsites, color=type, group=type), 
              method='loess', se=FALSE, span=1/50) + 
  scale_x_log10() + scale_y_log10() +
  facet_grid(Va ~ genlen) +
  ylab('fraction of segregating sites')  +
  xlab('frequency')




```

