---
title: "R sims neutral validation"
output: html_document
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=12, fig.height=8, 
               echo=TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(devtools)
library(tidyverse)
load_all()
```

## Variance Effective Population Size

First, we compare the variance effective population size, as calculated as $N_e
= 1/(2F)$ where $F=V(\Delta p)/p(1-p)$, to the real population sizes:

```{r}
load('../simdata/neutral-covs.Rdata')

Ne_ests <- neut_res %>% unnest(covs) %>% filter(t0 == t1) %>% 
  group_by(N, rho, t0, t1) %>% 
  summarize(var=mean(cov), Ne=1/(2*var*(1-1/(2*unique(N))))) %>%
  mutate(group=factor('A'))


Ne_ests %>%
  ggplot() + 
  geom_hline(aes(yintercept=N), color='red') + 
  geom_point(aes(t0, Ne)) + facet_grid(N ~ rho, scales='free_y') + 
  xlab('generation')

Ne_fits <- crossing(rho=unique(Ne_ests$rho), N=unique(Ne_ests$N), 
                    t0=0:50, group=factor("A")) %>% mutate(Ne=N)


# save data for figures
write_tsv(Ne_ests, '../data/supp_Ne_est_neutral.tsv')
devtools::use_data(Ne_ests, overwrite=TRUE)
devtools::use_data(Ne_fits, overwrite=TRUE)

neut_res %>% unnest(covs) %>% filter(t0 == t1) %>% 
  group_by(N, rho) %>% 
  summarize(var=mean(cov), Ne_corr=1/(2*var*(1-1/(2*unique(N)))),
            Ne=1/(2*var))


```

## LD Decay Rate

We simulate  a neutral Wright-Fisher population started with $N$ gametes with
111...111 alleles and $N$ gametes with 000...000 alleles, and measure $D$
between the two outermost markers. The initial state of the population creates a
baseline level of $D$ that we decay at different rates of $\rho$ and $N$
according to the simulations `r-sims/ld-test.r`. This should decay under a
finite population model as:

$$
D_t = D_1 (1-1/(2N))^{t-1} (1-r)^{t-1}
$$

We see this theoretic decay below in red compared to the average across
simulations in blue.

```{r, ld-decay}
load('../simdata/ld-res.Rdata')

D_decay <- function(D0, N, r, t, haldane=TRUE) {
  if (haldane)
    r <- haldane(r)
  D0*(1-1/(2*N))^t*(1-r)^t 
}

ldd <- ld_res %>% mutate(D=map(res, 'D')) %>% unnest(D) 
ldd_mean <- ldd %>% group_by(N, r, gen) %>% summarize(D=mean(D))
ldd_mean %>% ggplot(aes(gen, D)) + geom_point() + facet_grid(r ~ N)

# calculate D0 across replicates, and then calc the theoretic decay 
# through time using this for each N/r
D0 <- ldd %>% filter(gen==1) %>% summarize(D0 = mean(D)) %>% pull(D0)
ldd_fits  <- ldd %>% mutate(theory_D=pmap_dbl(list(D0, N, r, gen-1), D_decay)) 

ldd_fits %>%  ggplot() + geom_point(aes(gen, D), size=0.1, alpha=0.1) + 
  facet_grid(r ~ N) + 
  geom_line(data=ldd_mean, aes(gen, D)) + 
  geom_line(aes(gen, theory_D), col='red', size=1) + ylim(-0.1, 0.3) + ylab('D')


```

Now, we generate some data for supplementary figures:

```{r}

rparams <- c(0, 0.001, 0.01, 0.1, 0.5)
neut_ld_sims <- ldd  %>% arrange(r, N, gen) %>% 
  filter(r %in% rparams) %>%
  mutate(grp=factor('L'))  

neut_ld_fits <- ldd_fits %>% 
  filter(r %in% rparams) %>%
  mutate(D=theory_D, grp=factor('L')) %>%  arrange(r, N, gen)   %>%
  group_by(gen, r, N) %>%
  summarize(D=mean(D))

write_tsv(neut_ld_sims, '../data/supp_ld_neutral.tsv')
devtools::use_data(neut_ld_fits, overwrite=TRUE)
devtools::use_data(neut_ld_sims, overwrite=TRUE)



```

## Decay of Heterozygosity

First, we extract the neutral frequency and then calculate the mean site
heterozygosity. Then, we extract the initial heterozygosity at the start of the
simulations and use this to parameterize the decay of heterozygosity through time:

$$
H_t \approx H_1 (1-1/(2N))^t
$$

These match closely:

```{r}

freqdf <- function(mat) {
  setNames(melt(mat), c('gen', 'locus', 'freq'))
}

neut_freqs <- neut_res %>% mutate(neut_freqs=map(res, 'neut_freqs'), 
                           freqs=map(neut_freqs, freqdf)) %>% unnest(freqs)

neut_het <- neut_freqs %>% mutate(het=2*freq*(1-freq)) %>% 
              group_by(N, rho, gen) %>%
              summarize(het=mean(het)) 

neut_het_init <- neut_het %>% filter(gen==1) %>% select(h1=het)

neut_het_fit <- neut_het %>% left_join(neut_het_init) %>% mutate(ht=h1*(1-1/(2*N))^gen) 

neut_het_fit %>%
   ggplot(aes(gen, het)) + geom_point() + 
   geom_line(aes(gen, ht), color='red', size=1) + facet_grid(N~rho) +
   ylab('heterozygosity') + xlim('generation')

neut_het_fits <- neut_het_fit %>% mutate(het=ht) %>% select(-h1, -ht) %>% 
                   ungroup() %>% mutate(group=factor('A')) 

neut_het_sims <- neut_het %>% 
                   ungroup() %>% mutate(group=factor('A')) 

devtools::use_data(neut_het_fits, overwrite=TRUE)
devtools::use_data(neut_het_sims, overwrite=TRUE)




```
