---
title: "Single locus simulations"
output: html_document
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=12, fig.height=8, 
               echo=TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(devtools)
library(tidyverse)
load_all()
load('../simdata/sl-covs.Rdata') 

```

Parameterization notes: In my R sims, we use $w = 1 + \alpha g$ where $\alpha$
is the effect size and $g$ is the gene count for the selected allele. 


## Additive Genetic Variation in a Single Locus Model

There's something very subtle that happens with single-locus additive genetic
variation under our simulation routines. Because the selected site is chosen
randomly from a neutral stationary distribution, the mean allele frequency is
1/2, but the modes are near 0 and 1. The behavior of $p(1-p)$ over time at the
selected site is asymmetric: sites near fixation immediately fix and don't
contribute much to the mean trajectory of $p(1-p)$, where as sites at low
frequency raise in frequency with selection, increasing $V_A$ over the short
run, before then declining. Thus, $E(V_A(t))$ does not decrease; we can see
this by bootstrapping for $\alpha = 0$ and the free recombination case:

```{r}

sl_res %>% filter(alpha == 0.1, rec_frac > 0.4) %>% 
  unnest(stats) %>% filter(gen > 4) %>%
  modelr::bootstrap(20) %>%
  mutate(strap=map(strap, as_tibble)) %>% unnest(strap) %>%
  ggplot(aes(gen, wvar, group=.id)) + geom_smooth(se=FALSE)


```

Compare to conditioning on initial allele frequency:

```{r}

get_init_freq <- function(x) x$sel_freqs[1]


sl_wvar_trajs <- sl_res %>% filter(alpha == 0.1, rec_frac > 0.4) %>% 
  mutate(p0=map_dbl(res, get_init_freq)) %>%
  unnest(stats) %>% filter(gen > 4) %>%
  mutate(p0_bin=cut_number(p0, 5)) %>%
  #mutate(p0_bin=cut(p0, c(0, 0.05, 0.1, 0.3, 0.7, 0.9, 0.095, 1))) %>%
  group_by(p0_bin) %>%
  do(rs=modelr::bootstrap(., 20)) %>% unnest(rs) %>%
  mutate(strap=map(strap, as_tibble)) %>%
  unnest(strap) 


sl_wvar_trajs %>% ggplot(aes(gen, wvar, group=interaction(.id, p0_bin), color=p0_bin)) + 
  geom_smooth(se=FALSE) 

# write data for plot
devtools::use_data(sl_wvar_trajs, overwrite=TRUE)
write_tsv(sl_wvar_trajs, '../data/sl_wvar_trajs.tsv')


```

This explains why we see $V_A(t)$ has a hump above.

In our data, we want to condition on intermediate frequency. We do that below,
and calculate the theoretic single locus $V_A(t)$ (using `sl_va()`).


```{r}

# condition on intermediate freq
sl_res_int <- sl_res %>% select(-covs) %>% 
              mutate(init_freq=map_dbl(res, ~ .$sel_freqs[1])) %>% 
              filter(abs(init_freq - 0.5) < 0.1) 

# empirical variances
vark_int <- sl_res_int %>% 
          unnest(stats) %>% 
          group_by(N, alpha, rec_frac, gen) %>% 
          summarize(wvar=mean(wvar)) %>%
          ungroup() 

ngens <- max(vark_int$gen)


theory_vark <- 
  vark_int %>% select(N, alpha, rec_frac) %>% 
                 distinct(.keep_all=TRUE) %>%
                 mutate(va=map(alpha, ~ sl_var_traj(p0=0.5, ngens, s=.*2)))  %>%
                 unnest(va) %>% mutate(gen=gen+5)


all_vark_int <- vark_int %>% left_join(theory_vark) 

all_vark_int %>% 
  ggplot() + geom_point(aes(gen, wvar), color='red') + geom_line(aes(gen, theory_va), color='blue') + facet_grid(~alpha)


```

Compare to the case where we don't condition on intermediate frequency:


```{r}

# empirical variances
vark <- sl_res %>% 
          unnest(stats) %>% 
          group_by(N, alpha, rec_frac, gen) %>% 
          summarize(wvar=mean(wvar), kvar=mean(kvar), wbar=mean(wbar)) %>%
          ungroup() 

ngens <- max(vark$gen)
theory_vark <- 
  vark %>% select(N, alpha, rec_frac) %>% 
                 distinct(.keep_all=TRUE) %>%
                 mutate(va=map(alpha, ~ sl_var_traj(p0=0.5, ngens, s=.)))  %>%
                 unnest(va) %>% mutate(gen=gen)

all_vark <- vark %>% left_join(theory_vark) 


all_vark %>% mutate(kvar=(kvar-2)/2) %>% 
  mutate(rec_frac=as.factor(rec_frac)) %>%
  ggplot() + geom_point(aes(gen, wvar, color=rec_frac)) + geom_line(aes(gen+4, theory_va), color='blue', size=1.2) + facet_grid(~ alpha)

```

## Calculate Covariances

First we calculate the covariances. First, recall that there are two sites per
simulation: the selected site and the neutral site. We can calculate the
covariance by either calculating it on a single site as $\Delta p_t \Delta
p_s$, and then average the covariances across replicates. Or, we can bind each
of the neutral sites and calculate the covariance, calculating the covariance
by averaging across replicates (this essentially treats each replicate as a
different loci). The former is much, much nosier -- likely due to how fixations
are handled (and perhaps because the mean isn't subtracted off).  Here, we
reshape the neutral allele frequencies:


```{r}

sl_freqs <- sl_res %>% select(-covs, -stats) %>% 
              mutate(neut_freqs = map(res, get_neut_freqs)) %>%
              select(-res) %>%
              group_by(N, alpha, rec_frac) %>% nest() %>%
              # reshape neutral allele frequencies, binding 
              # replicates into a single matrix
              # that we can then calculate covariance over
              mutate(freqs=map(data, reshape_neut_freqs)) 

```

We calculate the temporal covariances on all of these replicates reshaped, such
that the covariance is averaging across replicates.

```{r}

sl_covs <- sl_freqs %>% 
  mutate(covs=map(freqs, ~ temp_cov(., as_df=TRUE, remove_fixed=FALSE))) %>% 
             unnest(covs)

pred <- all_vark %>% ungroup() %>% mutate(N=as.integer(as.character(N)), 
            cov=pmap_dbl(list(rec_frac, gen-5, theory_va, N), sl_analytic_cov))  %>%
           mutate(N=factor('1000'))

rec_params <- c(1e-4, 5e-4, 0.02)
alpha_params <- c(0.05, 0.1, 0.2)
covs5 <- sl_covs %>% filter(t1 > t0, t0 == 5, 
                            # $alpha %in% alpha_params,
                            #rec_frac %in% rec_params,
                            ) %>% mutate(gen=t1) %>%
           mutate(N=as.factor(N))


covs5 %>%
ggplot(aes(gen, cov)) + geom_point() + 
  geom_line(data=pred, aes(gen, cov), color='red') +
  geom_smooth() + 
  facet_grid(alpha~rec_frac, scales='free_y')

```

This shows that we get good fits, but at a certain point (high $\alpha$, low
$r$), we end up with interference.

We write the files necessary to produce the supplementary figure to file:


```{r}

sl_pd5f <- filter(covs5, alpha %in% alpha_params, rec_frac %in% rec_params) 
# only keep predictions when gen > 5, when selection is turned on. 
# our V_A just looks at sel. allele freq, not when sel. is turned 
# on so this is needed:
sl_predf <- filter(pred, alpha %in% alpha_params, rec_frac %in% rec_params, gen >= 5)


# tables and Rda files for plots:
write_tsv(sl_pd5f, '../data/sl_pd5f.tsv')
write_tsv(sl_predf, '../data/sl_predf.tsv')
devtools::use_data(sl_pd5f, overwrite=TRUE)
devtools::use_data(sl_predf, overwrite=TRUE)

panel_plot(sl_pd5f, sl_predf, gen, cov, alpha, rec_frac, N, pch_cex=1)

```



